Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study

### 2025-01-29 02_paper/02_study/02_expertise_paper/expertise_2025_01_23_MK.Rmd
x paper

# Step 1: Data Preparation
df_ttff_disrup <- df_aoi %>%
  filter(TOI %in% c(verbal_disruptions, physical_disruptions, lack_of_eagerness)) %>%
  dplyr::select(
    Participant,
    TOI,
    Time_to_first_fixation.Disruptive_Person
  ) %>%
  mutate(
    Group = ifelse(Participant < 200, "Novice", "Expert"),  # Assigning Group based on Participant
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) %>% # Keep only rows with valid categories
  drop_na() %>%
  filter(
    Time_to_first_fixation.Disruptive_Person > 0, # Exclude zero fixation times
    Time_to_first_fixation.Disruptive_Person <= 30000 # Cut-off at 30 seconds
  ) %>%
  mutate(
    Disrup_time_fixation_sec = round(Time_to_first_fixation.Disruptive_Person / 1000, 2)
  )

# Summarize the Disrup_time_fixation_sec by Participant
df_ttff_disrup_participant_summary <- df_ttff_disrup %>%
  group_by(Participant, Group) %>%
  summarise(
    Disrup_time_fixation_sec = round(mean(Disrup_time_fixation_sec, na.rm = TRUE), 2), # Average time to first fixation per participant
    .groups = "drop"
  )

@@

# Step 2: Descriptive Statistics by Group
ttff_disrup_table <- df_ttff_disrup_participant_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in sec" = round(mean(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "SD in sec" = round(sd(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Min in sec" = round(min(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Max in sec" = round(max(Disrup_time_fixation_sec, na.rm = TRUE), 2)
  )

knitr::kable(ttff_disrup_table,
             caption = "N, M, SD, Min & Max Time to First Fixation in Seconds (AOI Disruptive Student)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Step 3: Descriptive Statistics by Category and Group
summary_stats_category_group <- df_ttff_disrup %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n_distinct(Participant),  # Count unique participants
    Mean = round(mean(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    SD = round(sd(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Median = round(median(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Min = round(min(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    Max = round(max(Disrup_time_fixation_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Display the table
knitr::kable(
  summary_stats_category_group,
  caption = "Descriptive Statistics for Time to First Fixation by Category and Group",
  col.names = c("Category", "Group", "N (Participants)", "Mean", "SD", "Median", "IQR", "Min", "Max"), # Ensure the column names match
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Plot 1: Time to First Fixation by Group
plot_ttff_group <- ggplot(df_ttff_disrup, aes(x = Group, y = Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Time to First Fixation (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Time to First Fixation by Group") +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = max(df_ttff_disrup$Disrup_time_fixation_sec, na.rm = TRUE) * 1.1,
           label = paste0("Median = ", round(median(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"], na.rm = TRUE), 2),
                          "\nIQR = ", round(IQR(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"], na.rm = TRUE), 2)),
           hjust = 0.5, size = 4) +
  annotate("text", x = 2, y = max(df_ttff_disrup$Disrup_time_fixation_sec, na.rm = TRUE) * 1.1,
           label = paste0("Median = ", round(median(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"], na.rm = TRUE), 2),
                          "\nIQR = ", round(IQR(df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"], na.rm = TRUE), 2)),
           hjust = 0.5, size = 4)

# Display the first plot
plot_ttff_group

# Plot 2: Time to First Fixation by Group and Disruption Category
plot_ttff_category <- ggplot(df_ttff_disrup, aes(x = Group, y = Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "fixed", labeller = label_wrap_gen(width = 15)) +  # Use fixed scale for all facets  
  ylim(0, 30) + 
  labs(x = "Group", y = "Time to First Fixation (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Time to First Fixation by Group and Disruption Category") +
  theme(legend.position = "none")

# Display the second plot
plot_ttff_category

@@

# Perform the repeated-measures ANOVA with `Disrup_time_fixation_sec` as the dependent variable
anova_afex_ttff <- afex::aov_ez(
  id = "Participant",  # Random effect (each participant is measured multiple times)
  dv = "Disrup_time_fixation_sec",  # Dependent variable: Time to First Fixation (seconds)
  between = "Group",  # Between-subject factor: Group (Novice vs Expert)
  within = "Disruption_Category",  # Within-subject factor: Disruption Categories (Verbal, Physical, Lack of eagerness)
  data = df_ttff_disrup  # The data frame that contains the TTFF data
)

# Extract the ANOVA table
anova_results_ttff <- as.data.frame(anova_afex_ttff$anova_table)

# Recalculate generalized eta squared (η²) for effect size
eta_squared_results_ttff <- effectsize::eta_squared(anova_afex_ttff, partial = FALSE)

# Combine recalculated eta squared with the ANOVA table
anova_table_apa_ttff <- anova_results_ttff %>%
  rownames_to_column(var = "Effect") %>%  # Move row names to a column
  left_join(
    as.data.frame(eta_squared_results_ttff), 
    by = c("Effect" = "Parameter")  # Match the correct columns
  ) %>%
  mutate(
    `F` = round(`F`, 2),  # Round F-values
    `η²` = round(Eta2, 3),  # Round generalized eta squared
    `p-value` = ifelse(
      `Pr(>F)` < 0.05, "< .05", round(`Pr(>F)`, 2)  # Format p-values
    )
  ) %>%
  select(Effect, `num Df`, `den Df`, `F`, `p-value`, `η²`)  # Order columns, place η² last

# Display the ANOVA table for Time to First Fixation
knitr::kable(
  anova_table_apa_ttff,
  caption = "Repeated-Measures ANOVA Results for Time to First Fixation (Including Effect Sizes)",
  col.names = c("Effect", "Df1", "Df2", "F", "p", "η²"),
  digits = 2
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Perform post-hoc pairwise comparisons for Disruption_Category (within-subject factor)
post_hoc_disruption <- emmeans(anova_afex_ttff, pairwise ~ Disruption_Category)

@@

# Calculate Cohen's d for each pairwise comparison
d_LackOfEagerness_vs_Physical <- CohenD(
  x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Lack of eagerness to learn"],
  y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Physical disruptions"],
  na.rm = TRUE
)

d_LackOfEagerness_vs_Verbal <- CohenD(
  x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Lack of eagerness to learn"],
  y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Verbal disruptions"],
  na.rm = TRUE
)

d_Physical_vs_Verbal <- CohenD(
  x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Physical disruptions"],
  y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Verbal disruptions"],
  na.rm = TRUE
)

@@

@@

### Self-evaluation (8 classroom management items, 8 presence items)
```{r self-eval_8-8, echo=FALSE, results='asis'}

>>>>

# Step 1: Data Preparation
df_ttff_disrup <- df_aoi %>%
  filter(TOI %in% c(verbal_disruptions, physical_disruptions, lack_of_eagerness)) %>%
  dplyr::select(
    Participant,
    TOI,
    Time_to_first_fixation.Disruptive_Person
  ) %>%
  mutate(
    Group = ifelse(Participant < 200, "Novice", "Expert"),  # Assigning Group based on Participant
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) %>% # Keep only rows with valid categories
  drop_na() %>%
  filter(
    Time_to_first_fixation.Disruptive_Person > 0, # Exclude zero fixation times
    Time_to_first_fixation.Disruptive_Person <= 30000 # Cut-off at 30 seconds
  ) %>%
  mutate(
    Disrup_time_fixation_sec = round(Time_to_first_fixation.Disruptive_Person / 1000, 2),
    log_Disrup_time_fixation_sec = round(log(Disrup_time_fixation_sec), 2) # Log-transformed variable
  )

# Summarize the log-transformed variable by Participant
df_ttff_disrup_participant_summary <- df_ttff_disrup %>%
  group_by(Participant, Group) %>%
  summarise(
    log_Disrup_time_fixation_sec = round(mean(log_Disrup_time_fixation_sec, na.rm = TRUE), 2), # Average log time to first fixation per participant
    .groups = "drop"
  )

@@

# Step 2: Descriptive Statistics by Group
log_ttff_disrup_table <- df_ttff_disrup_participant_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M (log-sec)" = round(mean(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "SD (log-sec)" = round(sd(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Min (log-sec)" = round(min(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Max (log-sec)" = round(max(log_Disrup_time_fixation_sec, na.rm = TRUE), 2)
  )

knitr::kable(log_ttff_disrup_table,
             caption = "N, M, SD, Min & Max Log-Transformed Time to First Fixation (AOI Disruptive Student)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Step 3: Descriptive Statistics by Category and Group
summary_stats_log_category_group <- df_ttff_disrup %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n_distinct(Participant),  # Count unique participants
    "M (log-sec)" = round(mean(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "SD (log-sec)" = round(sd(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Min (log-sec)" = round(min(log_Disrup_time_fixation_sec, na.rm = TRUE), 2),
    "Max (log-sec)" = round(max(log_Disrup_time_fixation_sec, na.rm = TRUE), 2)
  )

# Display the table
knitr::kable(
  summary_stats_log_category_group,
  caption = "Descriptive Statistics for Log-Transformed Time to First Fixation by Category and Group",
  col.names = c("Category", "Group", "N (Participants)", "Mean", "SD", "Min", "Max"), # Ensure the column names match
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Plot 1: Log Time to First Fixation by Group
plot_log_ttff_group <- ggplot(df_ttff_disrup_participant_summary, aes(x = Group, y = log_Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Log Time to First Fixation (log-seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Log Time to First Fixation by Group") +
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) + # Manuelle Achseneinteilung
  theme(legend.position = "none")

# Display the first plot
plot_log_ttff_group

# Plot 2: Log Time to First Fixation by Group and Disruption Category
plot_log_ttff_category <- ggplot(df_ttff_disrup, aes(x = Group, y = log_Disrup_time_fixation_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "fixed", labeller = label_wrap_gen(width = 15)) +  # Use fixed scale for all facets  
  labs(x = "Group", y = "Log Time to First Fixation (log-seconds)") +
  theme_cowplot() +
  scale_y_continuous(limits = c(-5, 5), breaks = seq(-5, 5, 1)) + # Manuelle Achseneinteilung
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Log Time to First Fixation by Group and Disruption Category") +
  theme(legend.position = "none")

# Display the second plot
plot_log_ttff_category

@@

# Perform the repeated-measures ANOVA with `log_Disrup_time_fixation_sec` as the dependent variable
anova_afex_ttff_log <- afex::aov_ez(
  id = "Participant",  # Random effect (each participant is measured multiple times)
  dv = "log_Disrup_time_fixation_sec",  # Dependent variable: Log Time to First Fixation (log-seconds)
  between = "Group",  # Between-subject factor: Group (Novice vs Expert)
  within = "Disruption_Category",  # Within-subject factor: Disruption Categories (Verbal, Physical, Lack of eagerness)
  data = df_ttff_disrup  # The data frame that contains the log-transformed TTFF data
)

# Extract the ANOVA table
anova_results_ttff_log <- as.data.frame(anova_afex_ttff_log$anova_table)

# Recalculate generalized eta squared (η²) for effect size
eta_squared_results_ttff_log <- effectsize::eta_squared(anova_afex_ttff_log, partial = FALSE)

# Combine recalculated eta squared with the ANOVA table
anova_table_apa_ttff_log <- anova_results_ttff_log %>%
  rownames_to_column(var = "Effect") %>%  # Move row names to a column
  left_join(
    as.data.frame(eta_squared_results_ttff_log), 
    by = c("Effect" = "Parameter")  # Match the correct columns
  ) %>%
  mutate(
    `F` = round(`F`, 2),  # Round F-values
    `η²` = round(Eta2, 3),  # Round generalized eta squared
    `p-value` = ifelse(
      `Pr(>F)` < 0.05, "< .05", round(`Pr(>F)`, 2)  # Format p-values
    )
  ) %>%
  select(Effect, `num Df`, `den Df`, `F`, `p-value`, `η²`)  # Order columns, place η² last

# Display the ANOVA table for Log Time to First Fixation
knitr::kable(
  anova_table_apa_ttff_log,
  caption = "Repeated-Measures ANOVA Results for Log-Transformed Time to First Fixation (Including Effect Sizes)",
  col.names = c("Effect", "Df1", "Df2", "F", "p", "η²"),
  digits = 2
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Perform post-hoc pairwise comparisons for Disruption_Category (within-subject factor)
post_hoc_disruption <- emmeans(anova_afex_ttff_log, pairwise ~ Disruption_Category)

@@

# Calculate Cohen's d for each pairwise comparison
d_LackOfEagerness_vs_Physical <- CohenD(
  x = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Lack of eagerness to learn"],
  y = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Physical disruptions"],
  na.rm = TRUE
)

d_LackOfEagerness_vs_Verbal <- CohenD(
  x = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Lack of eagerness to learn"],
  y = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Verbal disruptions"],
  na.rm = TRUE
)

d_Physical_vs_Verbal <- CohenD(
  x = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Physical disruptions"],
  y = df_ttff_disrup$log_Disrup_time_fixation_sec[df_ttff_disrup$Disruption_Category == "Verbal disruptions"],
  na.rm = TRUE
)

@@

## Classroom Management Measures

@@

### self-evaluations of disruption prevention and management competencies (5 classroom management items + 3 self-developed, 8 presence items)
```{r self-eval_8-8, echo=FALSE, results='asis'}
